Jaki tytuł??

### 2020-05-20 article.Rmd
Add files via upload

---
title: "The imputation methods big test"
author: "Jakub Kosterna, Dawid Przybyliński & Hanna Zdulska"
date: "22/04/2020"
output:
  html_document:
    code_folding: hide
  pdf_document: default
---

@@

## Abstract

Choosing the best-suited imputation is the daily dilemma of every data scientist. Some of they believe the crux lies in the most advanced and sophisticated, the others trust the simplest of all possible.

In this document we will try to find one objectively the finest imputation method by testing five of popular ones on eight specially selected data sets prepared for these operations.

@@

We'll look at the two simple imputation methods with the following short markings:

1. *modeMedian* - missing data supplemented with medians from individual columns
2. *removeRows* - all the rows with any nonexistent values removed

```{r imputation_basic_functions, cache = TRUE}
imputation_mode_median <- function(df){

  Mode <- function(x) {
    ux <- unique(x)
    ux[which.max(tabulate(match(x, ux)))]
  }

  for (i in 1L:length(df)){
    if (sum(is.na(df[,i])) > 0){
      if (mode(df[,i]) == 'character' | is.factor(df[,i])){
        to_imp <- Mode(df[,i])
        df[,i][is.na(df[,i])] <- to_imp
      }
      else{
        to_imp <- median(df[,i], na.rm = TRUE) 
        df[,i][is.na(df[,i])] <- to_imp
      }
    }
  }

  return(df)
}
imputation_remove_rows <- function(df){
  return (na.omit(df))
}
```

... and also three more advanced algorithms from popular libraries:

3. *mice* - standard imputation with *maxit* parameter of value 0 and default 5 number of imputed datasets
4. *I4* - *vim* - default *kNN* implementation
5. *I5* - *missForest* - simple imputation offered by the package

```{r imputation_advanced_functions, message = FALSE, warning = FALSE, cache = TRUE}
library(mice)
imputation_fun_mice <- function(df){
  init <- mice(df, maxit=0) 
  meth <- init$method
  predM <- init$predictorMatrix
  imputed <- mice(df, method=meth, predictorMatrix=predM, m=5)
  completed <- complete(imputed)
  return(completed)
}
library(VIM)
imputation_fun_vim <- function(df){
  no_columns <- length(df)
  imputed <- kNN(df)
  imputed <- imputed[,1:no_columns]
  return(imputed)
}
library(missForest)
imputation_fun_missForest <- function(df){
  return(missForest(df)$ximp)
}

```

@@

```{r read_dataset, cache = TRUE}
DFT_REPO_DATASET_DIR = './dependencies/datasets'
read_dataset <- function(openml_id, dataset_dir = DFT_REPO_DATASET_DIR){

  if (!dir.exists(dataset_dir)){
    stop(paste(dataset_dir, 'does not exist' ))
  }

  dir <- paste(dataset_dir, paste('openml_dataset', openml_id, sep = '_'), sep ='/')
  if (!dir.exists(dir)){
    stop(paste(dir, 'does not exist' ))
  }

  start_dir <- getwd()

  # set right dir to code.R to acually work - it depends on dirlocation to create json
  setwd(dir)
  # use new env to avoid trashing globalenv
  surogate_env <- new.env(parent = .BaseNamespaceEnv)
  attach(surogate_env)
  source("code.R",surogate_env)

  j <- jsonlite::read_json('./dataset.json')
  j$dataset <- surogate_env$dataset
  setwd(start_dir)

  return(j)
}
```

In order to read all datasets we will use other function, which will return a dataframe containing all the important informations about the test matrices.

```{r read_all_datasets, cache = TRUE}
read_all_datasets <- function(dataset_dir = DFT_REPO_DATASET_DIR){

  if (!dir.exists(dataset_dir)){
    stop(paste(dataset_dir, 'does not exist'))
  }

  start_dir <- getwd()
  subdirs <- dir(dataset_dir)
  ids <- sapply(subdirs, function(dir){substr(dir, 16, nchar(dir))})
  datasets_combined <- lapply(ids, function(x){read_dataset(x, dataset_dir)})

  datasets_combined <- t(datasets_combined)
  return(unname(t(datasets_combined)))
}
```

All of datasets contain missing information denoted as NA and differ in size, number of features, number of observations and percentage of missing data. Table below decribes those informations.

```{r dataset description, message = FALSE, warning=FALSE, cache = TRUE}

dfs <- read_all_datasets()

ids <- lapply(dfs, function(x){x$id})
names <- lapply(dfs, function(x){x$data_name})
n_inst <- lapply(dfs, function(x){x$number_of_instances})
n_features <- lapply(dfs, function(x){x$number_of_features})
number_of_numeric_features <- lapply(dfs, function(x){x$number_of_numeric_features})
number_of_categorical_features <- lapply(dfs, function(x){x$number_of_categorical_features})
number_of_missings <- lapply(dfs, function(x){x$number_of_missings})
proc_missing <- lapply(dfs, function(x){x$number_of_missing/(x$number_of_instances * x$number_of_features)})
number_of_instances_with_missings <- lapply(dfs, function(x){x$number_of_instances_with_missings})
number_of_features_with_missings <- lapply(dfs, function(x){x$number_of_features_with_missings})

#fix wrong json
names[[10]] <- "okcupid-stem"

table <- data.frame("id" = unlist(ids), 
                    "name" = unlist(names),
                    "number_of_instances" = unlist(n_inst),
                    "number_of_features" = unlist(n_features),
                    "number_of_missing" = unlist(number_of_missings),
                    "procentage_of_missing" = round(unlist(proc_missing)*100,2))

knitr::kable(table)
```

@@

```{r train_test_split, cache = TRUE}
set.seed(1357)
train_test_split <- function(dataset, train_size){
  smp_size <- floor(train_size * nrow(dataset))
  typeof(smp_size)

  train_ind <- sample(seq_len(nrow(dataset)), size = smp_size)

  train <- dataset[train_ind, ]
  test <- dataset[-train_ind, ]

  return (list(train, test))
}
```

@@

```{r metrics, cache = TRUE}
get_confusion_matrix <- function(test, pred){
  return (table(Truth = test, Prediction = pred))
}
confusion_matrix_values <- function(confusion_matrix){
  TP <- confusion_matrix[2,2]
  TN <- confusion_matrix[1,1]
  FP <- confusion_matrix[1,2]
  FN <- confusion_matrix[2,1]
  return (c(TP, TN, FP, FN))
}
accuracy <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  return((conf_matrix[1] + conf_matrix[2]) / (conf_matrix[1] + conf_matrix[2] + conf_matrix[3] + conf_matrix[4]))
}
precision <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  return(conf_matrix[1]/ (conf_matrix[1] + conf_matrix[3]))
}
recall <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  return(conf_matrix[1] / (conf_matrix[1] + conf_matrix[4]))
}
f1 <- function(confusion_matrix){
  conf_matrix <- confusion_matrix_values(confusion_matrix)
  rec <- recall(confusion_matrix)
  prec <- precision(confusion_matrix)
  return(2 * (rec * prec) / (rec + prec))
}
```

@@

The results of our work will be checked using the implemented classification tree algorithm from the *rpart* package. We have chosen this algorithm because this classifier works well for almost any type of data - and for the most part numerical, and for the most part factors, it can handle and make a good model. The use of an algorithm using, for example, the distance from the nearest points or separating sets by clusters would not be recommended for each data frame, and given the diversity, here rpart should do a good job.

```{r get_result_function, warning = FALSE, message = FALSE, cache = TRUE}
library(rpart)
get_result <- function(dataset_list, imputation_fun){

dataset <- dataset_list$dataset
name_of_target <- dataset_list$target
# imputation
imputation_start = Sys.time() # start to measure time
imputated_dataset <- imputation_fun(dataset) 
imputation_stop = Sys.time() # end measuring time
# train test split
train_test <- train_test_split(imputated_dataset, 0.8)
train <- as.data.table(train_test[1])
test <- as.data.table(train_test[2])
# modelling
vars <- colnames(dataset)[colnames(dataset)!=name_of_target]
my_formula <- as.formula(paste(name_of_target, paste(vars, collapse=" + "), sep=" ~ "))
modelling_start = Sys.time() # start to measure time
tree_model <- rpart(formula = my_formula, data = train,
                    method = "class", control = rpart.control(cp = 0))
y_pred <- as.data.frame(predict(tree_model, test, type = "class"))
modelling_stop = Sys.time() # end measuring time
# calculating metrics
confusion_matrix <- get_confusion_matrix(test[[name_of_target]], y_pred[,1])
accuracy_v <- accuracy(confusion_matrix)
precision_v <- precision(confusion_matrix)
recall_v <- recall(confusion_matrix)
f1_v <- f1(confusion_matrix)
classification_report <- data.frame(accuracy_v, precision_v,
                                    recall_v, f1_v)
colnames(classification_report) <- c("accuracy", "precision",
                                     "recall", "f1")
dataset_list$dataset <- NULL
# in future maybe return all dataset_list ?
# for now stick with readability
imp_method_name <- deparse(substitute(imputation_fun))
return(list( dataset_id = dataset_list$id, 
             imp_method = imp_method_name,
             confusion_matrix = confusion_matrix,
             classification_report = classification_report,
             imputation_time = imputation_stop - imputation_start,
             modelling_time = modelling_stop - modelling_start))
}
data_all <- read_all_datasets()
# imputations and targets preparation
imputations <- list(imputation_fun_vim, imputation_fun_missForest,
                  imputation_remove_rows, imputation_mode_median, imputation_fun_mice)
targets <- lapply(data_all, function(d){d$target})
```

Here are the results. NA indicates that method returned error:

```{r warning=FALSE, cache = TRUE}
  library(knitr)
  results <- rep(list(0), 5)
  results[[1]] <- readRDS('./part_results/res1_new.rds')
  results[[2]] <- readRDS('./part_results/res2_new.rds')
  results[[3]] <- readRDS('./part_results/res3_new.rds')
  results[[4]] <- readRDS('./part_results/res4_new.rds')
  results[[5]] <- readRDS('./part_results/res5_new.rds')

  result_table <- function(ds_it){
    d <- data.frame(matrix(ncol = 7, nrow = 0))
    colnames(d) <- c("imputation", "accuracy", "precision", "recall", "f1", "imp_time", "mod_time")
    for (i in 1:length(imputations)){
      if (length(results[[i]][[ds_it]]) != 1){ # length equal 1 means "ERROR" was generated in partial results 
        d[i,] <- c(results[[i]][[ds_it]]$imp_method,
               round(as.numeric(results[[i]][[ds_it]]$classification_report$accuracy),3),   
               round(as.numeric(results[[i]][[ds_it]]$classification_report$precision),3),
               round(as.numeric(results[[i]][[ds_it]]$classification_report$recall),3),
               round(as.numeric(results[[i]][[ds_it]]$classification_report$f1),3),
               round(as.numeric(results[[i]][[ds_it]]$imputation_time),3),
               # @TODO konwersja jednostek, as.numeric nie rozroznia jednostek czasu
               round(as.numeric(results[[i]][[ds_it]]$modelling_time),3)) 
      }
      else{ # for error input imputation name
        if (i==3){
          d[i,1] <- "imputation_fun_mice" 
        }
        if (i==4){
          d[i,1] <- "imputation_fun_missForest" 
        }
        if (i==5){
          d[i,1] <- "imputation_fun_vim" 
        }
      }
    }
    d
  }

  # change i bound to number of datasets 
  #for (i in 1:length(results[[1]])){ # won't use it, because the output looks bad
  #  kable(result_table(i), caption = paste("Dataset id: ", results[[2]][[i]]$dataset_id, sep = ""))
  #}

  kable(result_table(1), caption = paste("Dataset id: ", results[[2]][[1]]$dataset_id, sep = ""))
  kable(result_table(2), caption = paste("Dataset id: ", results[[2]][[2]]$dataset_id, sep = ""))
  kable(result_table(3), caption = paste("Dataset id: ", results[[2]][[3]]$dataset_id, sep = ""))
  kable(result_table(4), caption = paste("Dataset id: ", results[[2]][[4]]$dataset_id, sep = ""))
  kable(result_table(5), caption = paste("Dataset id: ", results[[2]][[5]]$dataset_id, sep = ""))
  kable(result_table(6), caption = paste("Dataset id: ", results[[2]][[6]]$dataset_id, sep = ""))
  kable(result_table(7), caption = paste("Dataset id: ", results[[2]][[7]]$dataset_id, sep = ""))
  kable(result_table(8), caption = paste("Dataset id: ", results[[2]][[8]]$dataset_id, sep = ""))
  kable(result_table(9), caption = paste("Dataset id: ", results[[2]][[9]]$dataset_id, sep = ""))
  kable(result_table(10), caption = paste("Dataset id: ", results[[2]][[10]]$dataset_id, sep = ""))
  kable(result_table(11), caption = paste("Dataset id: ", results[[2]][[11]]$dataset_id, sep = ""))
  kable(result_table(12), caption = paste("Dataset id: ", results[[2]][[12]]$dataset_id, sep = ""))
  kable(result_table(13), caption = paste("Dataset id: ", results[[2]][[13]]$dataset_id, sep = ""))
  kable(result_table(14), caption = paste("Dataset id: ", results[[2]][[14]]$dataset_id, sep = ""))
```

@@

```{r imputation_times_counting, cache = TRUE}
ndatasets <- length(results[[1]])
nimputations <- length(imputations)
imputation_names <- c("remove", "median", "mice", "missForest", "VIM")
dataset_ids <- rep(NULL, ndatasets)
for(i in 1:ndatasets){
  dataset_ids[i] <- results[[2]][[i]]$dataset_id
}

imputation_times <- data.frame(rep(0, ndatasets))
for (imputation_id in 1:nimputations){
  imp_id_times <- rep(NULL, ndatasets)

  for(dataset_id in 1:ndatasets){
    if("imputation_time" %in% names(results[[imputation_id]][[dataset_id]])){
      imp_time <- results[[imputation_id]][[dataset_id]]$imputation_time
      imp_time <- as.numeric(imp_time, units="secs")
      imp_id_times[dataset_id] <- imp_time
    }
  }
  imputation_times <- cbind(imputation_times, imp_id_times)
}

imputation_times[1] <- dataset_ids
colnames(imputation_times) <- c("dataset_id", imputation_names)
imputation_times$dataset_id <- as.factor(imputation_times$dataset_id)
knitr::kable(imputation_times)
```

@@

```{r imputation_times_plot, message = FALSE, warning = FALSE, cache = TRUE}
library(ggplot2)
# install.packages("reshape") # if not installed
library(reshape)

# in order to connect colors and shapes in legend see:
# https://stackoverflow.com/questions/12410908/combine-legends-for-color-and-shape-into-a-single-legend

colors <- c("remove" = "orange", "median" = "cyan", "mice" = "green",
            "missForest" = "red", "VIM" = "blue")
imptimes <- melt(imputation_times)

require(scales)
ggplot(data=imptimes, aes(x = dataset_id, y = value, color = variable, shape = variable)) +
  geom_point() + 
  ggtitle("Datasets' imputations times") +
  theme_gray() +
  scale_y_continuous(trans = log2_trans(),
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))) +
  labs(x = "Dataset Id",
       y = "Imputation time [in seconds]",
       color = "Legend") +
  scale_color_manual(name = "imputation name", values = colors) +
  theme(legend.position="right",
        plot.title = element_text(hjust = 0.5)) +
  #scale_colour_manual(name = "Treatment & State",
  #                    labels = c("Control, Non-F", "Control, Flwr", "Exclosure, Non-F", "Exclosure, Flwr"),
  #                    values = c("blue", "red", "blue", "red")) +   
  scale_shape_manual(name = "imputation name",
                     values = 15:19L)

```

@@

```{r imputation_times_measures, cache = TRUE}
means <- rep(NA, nimputations)
medians <- rep(NA, nimputations)
stDeviations <- rep(NA, nimputations)

for(i in (1:nimputations)){
  means[i] <- mean(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
  means[i] <- round(means[i], 3)
  medians[i] <- median(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
  medians[i] <- round(medians[i], 3)
  stDeviations[i] <- sd(imputation_times[imputation_names[i]][!is.na(imputation_times[imputation_names[i]])])
  stDeviations[i] <- round(stDeviations[i], 3)
}

imputation_times_measures <- data.frame(imputation_names, means, medians, stDeviations)
knitr::kable(imputation_times_measures)
```

@@

```{r imputation_times_measures_little, cache = TRUE}
dataset_ids_fullimp <- c(2, 5, 8, 11, 13, 14)
ndatasets_fullimp <- length(dataset_ids_fullimp)

means <- rep(NA, nimputations)
medians <- rep(NA, nimputations)
stDeviations <- rep(NA, nimputations)

for(i in 1:nimputations){
  means[i] <- mean(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
  means[i] <- round(means[i], 3)
  medians[i] <- median(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
  medians[i] <- round(medians[i], 3)
  stDeviations[i] <- sd(imputation_times[imputation_names[i]][dataset_ids_fullimp,])
  stDeviations[i] <- round(stDeviations[i], 3)
}

imputation_times_measures_fullimp <- data.frame(imputation_names, means, medians, stDeviations)
knitr::kable(imputation_times_measures_fullimp)
```

@@

## Comparing modeling times

Let's also take a look at modeling times.

```{r modeling_times_counting, cache = TRUE}
modeling_times <- data.frame(rep(0, ndatasets))

for (imputation_id in 1:nimputations){
  mod_id_times <- rep(NULL, ndatasets)

  for(dataset_id in 1:ndatasets){
    if("imputation_time" %in% names(results[[imputation_id]][[dataset_id]])){
      mod_time <- results[[imputation_id]][[dataset_id]]$modelling_time
      mod_time <- as.numeric(mod_time, units="secs")
      mod_id_times[dataset_id] <- mod_time
    }
  }
  modeling_times <- cbind(modeling_times, mod_id_times)
}

modeling_times[1] <- dataset_ids
colnames(modeling_times) <- c("dataset_id", imputation_names)
imputation_times$dataset_id <- as.factor(modeling_times$dataset_id)
knitr::kable(modeling_times)
```

Very interesting! The longest modeling time took one of the largest sets... on deficiencies supplemented by median. And it took apparently longer than other modeling! could a random tree have more to do with a lot of the same values? Probably yes. Let's visualize our results.

```{r modeling_times_plot, message = FALSE, warning = FALSE, cache = TRUE}
modeling_times$dataset_id <- as.factor(modeling_times$dataset_id)
modtimes <- melt(modeling_times)

require(scales)
ggplot(data=modtimes, aes(x = dataset_id, y = value, color = variable, shape = variable)) +
  geom_point() + 
  ggtitle("Datasets' modeling times") +
  theme_light() +
  scale_y_continuous(trans = log2_trans(),
    breaks = trans_breaks("log2", function(x) 2^x),
    labels = trans_format("log2", math_format(2^.x))) +
  labs(x = "Dataset Id",
       y = "Modeling time [in seconds]",
       color = "Legend") +
  scale_color_manual(name = "imputation name", values = colors) +
  theme(legend.position="right",
        plot.title = element_text(hjust = 0.5))
```

Here the results seem to be way more scattered. Let's check on measures also.

```{r modeling_times_measures, cache = TRUE}
means <- rep(NA, nimputations)
medians <- rep(NA, nimputations)
stDeviations <- rep(NA, nimputations)

for(i in (1:nimputations)){
  means[i] <- mean(modeling_times[imputation_names[i]][!is.na(modeling_times[imputation_names[i]])])
  means[i] <- round(means[i], 3)
  medians[i] <- median(modeling_times[imputation_names[i]][!is.na(modeling_times[imputation_names[i]])])
  medians[i] <- round(medians[i], 3)
  stDeviations[i] <- sd(modeling_times[imputation_names[i]][!is.na(modeling_times[imputation_names[i]])])
  stDeviations[i] <- round(stDeviations[i], 3)
}

modeling_times_measures <- data.frame(imputation_names, means, medians, stDeviations)
knitr::kable(modeling_times_measures)
```

The powerful but also fast tree classifier from *rpart* (both looking at the median and average) counted everything in less than a second. The largest fluctuations can be observed for the method of *median* and *VIM* supplementation - for them the standard deviation is less than two and a half seconds. You can also see very similar all three average measures for these two, probably due to the operation of *VIM* as a result giving similar effects as the *median*. Interestingly, unbeatably the fastest tree was built for the *mice* algorithm ... but wait! Maybe we'd better compare only the imputations for those sets for which we've done everything.

```{r modeling_times_measures_little, cache = TRUE}
means <- rep(NA, nimputations)
medians <- rep(NA, nimputations)
stDeviations <- rep(NA, nimputations)

for(i in 1:nimputations){
  means[i] <- mean(modeling_times[imputation_names[i]][dataset_ids_fullimp,])
  means[i] <- round(means[i], 3)
  medians[i] <- median(modeling_times[imputation_names[i]][dataset_ids_fullimp,])
  medians[i] <- round(medians[i], 3)
  stDeviations[i] <- sd(modeling_times[imputation_names[i]][dataset_ids_fullimp,])
  stDeviations[i] <- round(stDeviations[i], 3)
}

modeling_times_measures_fullimp <- data.frame(imputation_names, means, medians, stDeviations)
knitr::kable(modeling_times_measures_fullimp)
```

On the attached non-manipulated table, all average measures are almost identical. The conclusion is that the method of imputation probably has almost no effect on the modeling time - although it is difficult to say, given that we had a fairly small number of sets to compare and we have values in the order of hundredths of a second - probably just the operation of the computer pretty badly distorts the result.

@@

```{r measures_tables, cache = TRUE}
bin_ids <- c(1, 3, 4, 5, 6, 8, 9, 10, 11, 12, 13, 14)

accuracies <- data.frame(rep(0, ndatasets))
precisions <- data.frame(rep(0, ndatasets))
recalls <- data.frame(rep(0, ndatasets))
f1s <- data.frame(rep(0, ndatasets))

for (imputation_id in 1:nimputations){
  acc_id <- rep(NA, ndatasets)
  pre_id <- rep(NA, ndatasets)
  rec_id <- rep(NA, ndatasets)
  f1_id <- rep(NA, ndatasets)

  for(dataset_id in bin_ids){
    if("classification_report" %in% names(results[[imputation_id]][[dataset_id]])){
      accuracy <- results[[imputation_id]][[dataset_id]]$classification_report$accuracy
      acc_id[dataset_id] <- accuracy
      precision <- results[[imputation_id]][[dataset_id]]$classification_report$precision
      pre_id[dataset_id] <- precision
      recall <- results[[imputation_id]][[dataset_id]]$classification_report$recall
      rec_id[dataset_id] <- recall
      f1 <- results[[imputation_id]][[dataset_id]]$classification_report$f1
      f1_id[dataset_id] <- f1
    }
  }
  accuracies <- cbind(accuracies, acc_id)
  precisions <- cbind(precisions, pre_id)
  recalls <- cbind(recalls, rec_id)
  f1s <- cbind(f1s, f1_id)
}

accuracies[1] <- dataset_ids
colnames(accuracies) <- c("dataset_id", imputation_names)
accuracies$dataset_id <- as.factor(accuracies$dataset_id)
accuracies <- accuracies[-c(2, 7), ] # delete  non-binary classification datasets
knitr::kable(accuracies, caption = "Accuracies")

precisions[1] <- dataset_ids
colnames(precisions) <- c("dataset_id", imputation_names)
precisions$dataset_id <- as.factor(precisions$dataset_id)
precisions <- precisions[-c(2, 7), ] # delete  non-binary classification datasets
knitr::kable(precisions, caption = "Precisions")

recalls[1] <- dataset_ids
colnames(recalls) <- c("dataset_id", imputation_names)
recalls$dataset_id <- as.factor(recalls$dataset_id)
recalls <- recalls[-c(2, 7), ] # delete  non-binary classification datasets
knitr::kable(recalls, caption = "Recalls")

f1s[1] <- dataset_ids
colnames(f1s) <- c("dataset_id", imputation_names)
f1s$dataset_id <- as.factor(f1s$dataset_id)
f1s <- f1s[-c(2, 7), ] # delete  non-binary classification datasets
knitr::kable(f1s, caption = "F1s")
```

@@

```{r measures_visualizations, message = FALSE, warning = FALSE, cache = TRUE}
accMelt <- melt(accuracies)
accPlot <- ggplot(data=accMelt, aes(x = dataset_id, y = value, color = variable)) +
  geom_point() + 
  ggtitle("Datasets' accuracies") +
  ylim(0, 1) +
  labs(x = "Dataset Id",
       y = "Accuracy",
       color = "Legend") +
  scale_color_manual(name = "accuracy", values = colors) +
  theme(legend.position="right",
        plot.title = element_text(hjust = 0.5))
accPlot

preMelt <- melt(precisions)
prePlot <- ggplot(data=preMelt, aes(x = dataset_id, y = value, color = variable)) +
  geom_point() + 
  ggtitle("Datasets' precisions") +
  ylim(0, 1) +
  labs(x = "Dataset Id",
       y = "Precision",
       color = "Legend") +
  scale_color_manual(name = "precision", values = colors) +
  theme(legend.position="right",
        plot.title = element_text(hjust = 0.5))
prePlot

recMelt <- melt(recalls)
recPlot <- ggplot(data=recMelt, aes(x = dataset_id, y = value, color = variable)) +
  geom_point() + 
  ggtitle("Datasets' recalls") +
  ylim(0, 1) +
  labs(x = "Dataset Id",
       y = "Recall",
       color = "Legend") +
  scale_color_manual(name = "recall", values = colors) +
  theme(legend.position="right",
        plot.title = element_text(hjust = 0.5))
recPlot

f1Melt <- melt(f1s)
f1Plot <- ggplot(data=f1Melt, aes(x = dataset_id, y = value, color = variable)) +
  geom_point() + 
  ggtitle("Datasets' f1s") +
  ylim(0, 1) +
  labs(x = "Dataset Id",
       y = "F1",
       color = "Legend") +
  scale_color_manual(name = "f1", values = colors) +
  theme(legend.position="right",
        plot.title = element_text(hjust = 0.5))
f1Plot

library(gridExtra)
grid.arrange(accPlot +
               ggtitle("Accuracies") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
              prePlot +
               ggtitle("Precisions") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
              recPlot +
               ggtitle("Recalls") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
             f1Plot +
               ggtitle("F1s") +
               theme(legend.position = "none",
                axis.line.x = element_blank(),
                axis.ticks.x = element_blank(),
                axis.text.x = element_blank(),
                axis.title = element_blank()),
             nrow = 2,
             top = "Measures values for different datasets' imputations")
```

@@

```{r measures_measures, cache = TRUE}
accMeans <- rep(NA, nimputations)
accMedians <- rep(NA, nimputations)
accStDeviations <- rep(NA, nimputations)

preMeans <- rep(NA, nimputations)
preMedians <- rep(NA, nimputations)
preStDeviations <- rep(NA, nimputations)

recMeans <- rep(NA, nimputations)
recMedians <- rep(NA, nimputations)
recStDeviations <- rep(NA, nimputations)

f1sMeans <- rep(NA, nimputations)
f1sMedians <- rep(NA, nimputations)
f1sStDeviations <- rep(NA, nimputations)

for(i in (1:nimputations)){
    accMeans[i] <- mean(accuracies[imputation_names[i]][!is.na(accuracies[imputation_names[i]])])
    accMeans[i] <- round(accMeans[i], 3)
    accMedians[i] <- median(accuracies[imputation_names[i]][!is.na(accuracies[imputation_names[i]])])
    accMedians[i] <- round(accMedians[i], 3)
    accStDeviations[i] <- sd(accuracies[imputation_names[i]][!is.na(accuracies[imputation_names[i]])])
    accStDeviations[i] <- round(stDeviations[i], 3)

    preMeans[i] <- mean(precisions[imputation_names[i]][!is.na(precisions[imputation_names[i]])])
    preMeans[i] <- round(preMeans[i], 3)
    preMedians[i] <- median(precisions[imputation_names[i]][!is.na(precisions[imputation_names[i]])])
    preMedians[i] <- round(preMedians[i], 3)
    preStDeviations[i] <- sd(precisions[imputation_names[i]][!is.na(precisions[imputation_names[i]])])
    preStDeviations[i] <- round(preStDeviations[i], 3)

    recMeans[i] <- mean(recalls[imputation_names[i]][!is.na(recalls[imputation_names[i]])])
    recMeans[i] <- round(recMeans[i], 3)
    recMedians[i] <- median(recalls[imputation_names[i]][!is.na(recalls[imputation_names[i]])])
    recMedians[i] <- round(recMedians[i], 3)
    recStDeviations[i] <- sd(recalls[imputation_names[i]][!is.na(recalls[imputation_names[i]])])
    recStDeviations[i] <- round(recStDeviations[i], 3)

    f1sMeans[i] <- mean(f1s[imputation_names[i]][!is.na(f1s[imputation_names[i]])])
    f1sMeans[i] <- round(f1sMeans[i], 3)
    f1sMedians[i] <- median(f1s[imputation_names[i]][!is.na(f1s[imputation_names[i]])])
    f1sMedians[i] <- round(f1sMedians[i], 3)
    f1sStDeviations[i] <- sd(f1s[imputation_names[i]][!is.na(f1s[imputation_names[i]])])
    f1sStDeviations[i] <- round(f1sStDeviations[i], 3)
}

measures_measures <- data.frame(imputation_names, accMeans, accMedians, accStDeviations,
                                preMeans, preMedians, preStDeviations,
                                recMeans, recMedians, recStDeviations,
                                f1sMeans, f1sMedians, f1sStDeviations)

knitr::kable(measures_measures, content = "Model measures' concentrations' measures")
```

@@

```{r correct_measures_measures, cache = TRUE}
ids_all_model_measures <- c(4, 6, 9, 11, 12)

accMeans <- rep(NA, nimputations)
accMedians <- rep(NA, nimputations)
accStDeviations <- rep(NA, nimputations)

preMeans <- rep(NA, nimputations)
preMedians <- rep(NA, nimputations)
preStDeviations <- rep(NA, nimputations)

recMeans <- rep(NA, nimputations)
recMedians <- rep(NA, nimputations)
recStDeviations <- rep(NA, nimputations)

f1sMeans <- rep(NA, nimputations)
f1sMedians <- rep(NA, nimputations)
f1sStDeviations <- rep(NA, nimputations)

for(i in 1:nimputations){
    accMeans[i] <- mean(accuracies[imputation_names[i]][ids_all_model_measures,])
    accMeans[i] <- round(accMeans[i], 3)
    accMedians[i] <- median(accuracies[imputation_names[i]][ids_all_model_measures,])
    accMedians[i] <- round(accMedians[i], 3)
    accStDeviations[i] <- sd(accuracies[imputation_names[i]][ids_all_model_measures,])
    accStDeviations[i] <- round(stDeviations[i], 3)

    preMeans[i] <- mean(precisions[imputation_names[i]][ids_all_model_measures,])
    preMeans[i] <- round(preMeans[i], 3)
    preMedians[i] <- median(precisions[imputation_names[i]][ids_all_model_measures,])
    preMedians[i] <- round(preMedians[i], 3)
    preStDeviations[i] <- sd(precisions[imputation_names[i]][ids_all_model_measures,])
    preStDeviations[i] <- round(preStDeviations[i], 3)

    recMeans[i] <- mean(recalls[imputation_names[i]][ids_all_model_measures,])
    recMeans[i] <- round(recMeans[i], 3)
    recMedians[i] <- median(recalls[imputation_names[i]][ids_all_model_measures,])
    recMedians[i] <- round(recMedians[i], 3)
    recStDeviations[i] <- sd(recalls[imputation_names[i]][ids_all_model_measures,])
    recStDeviations[i] <- round(recStDeviations[i], 3)

    f1sMeans[i] <- mean(f1s[imputation_names[i]][ids_all_model_measures,])
    f1sMeans[i] <- round(f1sMeans[i], 3)
    f1sMedians[i] <- median(f1s[imputation_names[i]][ids_all_model_measures,])
    f1sMedians[i] <- round(f1sMedians[i], 3)
    f1sStDeviations[i] <- sd(f1s[imputation_names[i]][ids_all_model_measures,])
    f1sStDeviations[i] <- round(f1sStDeviations[i], 3)
}

correct_measures_measures <- data.frame(imputation_names, accMeans, accMedians, accStDeviations,
                                preMeans, preMedians, preStDeviations,
                                recMeans, recMedians, recStDeviations,
                                f1sMeans, f1sMedians, f1sStDeviations)

knitr::kable(correct_measures_measures, content = "CORRECT model measures' concentrations' measures")
```

@@

```{r model_measures_plots, message = FALSE, cache = TRUE}
colors2 <- c("mean" = "blue", "median" = "orange")

>>>>

---
title: "Jaki tytuł??"
author: "Jakub Kosterna, Dawid Przybyliński & Hanna Zdulska"
date: "20/05/2020"
output:
  html_document:
    code_folding: hide
  pdf_document: default
---

@@

# Abstract

Choosing the best-suited imputation is the daily dilemma of every data scientist. Some of them believe the crux lies in the most advanced and sophisticated, the others trust the simplest of all possible.

In this document we will try to find the one, objectively finest imputation method by testing six of popular ones on forteen specially selected data sets prepared for these operations.

# Introduction + motivation

@@

We'll look at the three simple imputation methods and three more advanced algorithms offered by popular packages.

TODO: czemu one; lanie wody o tym, że są fajne bo różne i badamy różne spojrzenia

1. *mode / median replace* - basic method which puts in a place of missing data median cells from their columns for numerical values and dominants for categorical ones.
2. *remove rows* - trivial solution of removing rows containing any missing data.
3. *bogo replace* - simple algorithm replacing NA values with random numbers or factors from their columns.
4. *mice* ^1^ - method creating multiple imputations for multivariate missing data, based on Fully Conditional Specification, where each incomplete variable is imputed by a separate model. Standard imputation with *maxit* parameter of value 0 and default 5 number of imputed datasets
5. *VIM* ^2^ - standard k-nearest neighbors algorithm taken from the library
6. *missForest* ^3^ - simple imputation based on random forest offered by the package

## Datasets

TODO: ładniejszy format

@@

TODO: tylko na sześciu wszystko się udało

## Binary classification algorithms

TODO: pięć metod, ich krótki opis i również krótkie stwierdzenie, że sa fajne bo różne i wgl.

# Description of methodology and results

In order to read all datasets we used another function, which returned a dataframe containing all the important informations about the test matrices.

TODO: trochę o metodzie, o funkcjach czytających; PARALLEL, itd

@@

@@

TODO: MCC jeszcze, czemu f1

@@

The results of our work will be checked using the implemented classification tree algorithm from the *rpart ^4^* package. We have chosen this algorithm because this classifier works well for almost any type of data - and for the most part numerical, and for the most part factors, it can handle and make a decent model. The use of an algorithm using, for example, the distance from the nearest points or separating sets by clusters would not be recommended for each data frame, and given the diversity, here rpart should do a good job.

Also, it is worth mentioning that removing rows computes a different dataset compared to other techniques. Because of that one needs to be cautious while analysing and comparing remove_rows to other imputation methods. For one's own research it is probably worth replacing removing rows with removing columns for datasets with few features containing missing values. 
Here are the results. NA indicates that method returned error:

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@
